Further Evidence of Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution

### 2022-03-22 src/ms.tex
revising cmd

\begin{figure*}
    \includegraphics[width=\linewidth]{./figures/cmd.pdf}
    \caption{Color-magnitude diagram. In each panel, a 2-d histogram of all Kepler targets is shown in grayscale, with lighter shades representing a higher number of targets in each cell. In the middle and right panels, the \citet{Santos2021} targets with photometric rotation period detections are overplotted in orange. In the right panel, the CKS long-period pile-up stars identified in this work are shown by blue points and a Gaussian kernel density estimation of the \citet{Hall2021} asteroseismic sample is shown by blue contours.}
    \label{fig:cmd}
\end{figure*}

\subsection{Where do the pile-ups end?} \label{subsec:extent}
It appears from Figures~\ref{fig:kde} and \ref{fig:models} that the number density of stars on the \edit1{long-period pile-up} declines towards cooler \teff, as predicted by the WMB model \citep[see Figure 13 of][]{vanSaders2019}. \edit1{The number density of stars on the short-period pile-up similarly declines towards cooler \teff. If the short-period pile-up is due to core–envelope coupling, one might expect an opposite trend of increasing number density toward cooler \teff, since the core–envelope coupling timescales and hence the ``stalled'' braking phases are longer for lower-mass stars \citep{Curtis2020}. However, the observed \teff--\prot distribution will also depend sensitively on the selection functions and observational biases inherent to both Kepler and the source of \teff (e.g. LAMOST).} While it not clear whether or not these declines are astrophysical in nature, the result of \edit1{the selection functions or observational biases inherent to Kepler or LAMOST}, or some combination of effects, we attempted to characterize the extent of the pile-ups through the following approach. We found through inspection that constant Rossby curves of Ro=0.5 and Ro=1.3 (for the short- and long-period pile-ups, respectively) appear to describe the highest density contours found from Gaussian kernel density estimation of the \lamostmcq sample. In windows of 10~K width we measured the fraction of stars with periods within 1~d of each of the two constant Rossby curves, relative to the total number of stars in that \teff window. We found that the relative fraction of stars on the long-period pile-up declines rapidly between 6500~K and 5800~K, by more than half in that temperature range before plateauing at cooler \teff (Figure~\ref{fig:fraction}). The relative fraction of stars on the short-period \edit1{pile-up} declines more slowly, and below the temperature of the Sun, the relative fractions of stars on the two pile-ups are nearly equal. Thus, it is not yet clear if either pile-up extends to temperatures cooler than \tsun.

>>>>

\begin{figure*}
    \includegraphics[width=\linewidth]{./figures/cmd.pdf}
    \caption{Gaia DR2 color-magnitude diagram. In each panel, a 2-d histogram of all Kepler targets is shown in grayscale, with lighter shades representing a higher number of targets in each cell. In the middle and right panels, the \citet{Santos2021} targets with photometric rotation period detections are overplotted in orange. In the right panel, the CKS long-period pile-up stars identified in this work are shown by blue points and a Gaussian kernel density estimation of the \citet{Hall2021} asteroseismic sample is shown by blue contours.}
    \label{fig:cmd}
\end{figure*}

\subsection{Where do the pile-ups end?} \label{subsec:extent}
It appears from Figures~\ref{fig:kde} and \ref{fig:models} that the number density of stars on the \edit1{long-period pile-up} declines towards cooler \teff, as predicted by the WMB model \citep[see Figure 13 of][]{vanSaders2019}. \edit1{The number density of stars on the short-period pile-up similarly declines towards cooler \teff. If the short-period pile-up is due to core–envelope coupling, one might expect an opposite trend of increasing number density toward cooler \teff, since the core–envelope coupling timescales and hence the ``stalled'' braking phases are longer for lower-mass stars \citep{Curtis2020}. However, the observed \teff--\prot distribution will also depend sensitively on the selection functions and observational biases inherent to both Kepler and the source of \teff (e.g. LAMOST).}

While it not clear whether or not these declines are astrophysical in nature, the result of \edit1{the selection functions or observational biases inherent to Kepler or LAMOST}, or some combination of effects, we attempted to characterize the extent of the pile-ups through the following approach. We found through inspection that constant Rossby curves of Ro=0.5 and Ro=1.3 (for the short- and long-period pile-ups, respectively) appear to describe the highest density contours found from Gaussian kernel density estimation of the \lamostmcq sample. In windows of 10~K width we measured the fraction of stars with periods within 1~d of each of the two constant Rossby curves, relative to the total number of stars in that \teff window. We found that the relative fraction of stars on the long-period pile-up declines rapidly between 6500~K and 5800~K, by more than half in that temperature range before plateauing at cooler \teff (Figure~\ref{fig:fraction}). The relative fraction of stars on the short-period \edit1{pile-up} declines more slowly, and below the temperature of the Sun, the relative fractions of stars on the two pile-ups are nearly equal. Thus, it is not yet clear if either pile-up extends to temperatures cooler than \tsun.
